---
title: "Logistic Regression Test Uptake"
author: "Olaide Adebayo-Clement"
date: "2024-05-20"
output: word_document
---

```{r setup, include=FALSE}
rm(list = ls(all.names = TRUE))

knitr::opts_chunk$set(
  echo = FALSE,
  error = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 16.00,
  fig.height = 13.00,
  fig.align = 'center',
  fig.topcaption = TRUE,
  out.width = '100%',
  out.height = 'auto'
)

if (!requireNamespace("pacman", quietly = TRUE)) install.packages("pacman")

pacman::p_load(
  tidyverse, readxl, DBI, dplyr, odbc, janitor, gtsummary,
  scales, gt, purrr, broom, here, readr, car, MASS, mice,
  forestplot, stringr, tidyr, ggplot2, grid, gridExtra,
  ggpubr, stats, parallel, data.table, forcats, epiR,
  rpart, conflicted, doparallel, foreach, openxlsx,
  caret, pROC, lme4, glmmTMB, broom.mixed, geepack
)

conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::recode)
```

```{r database-connection}
db_conn <- odbc::dbConnect(
  odbc::odbc(),
  .connection_string = "driver={SQL Server};server=DB_SERVER;database=ANALYSIS_DB;Encrypt=true;trusted_connection=true",
  timeout = 60,
  timezone = Sys.timezone(),
  timezone_out = Sys.timezone()
)
```

```{r data-preparation}
load_service_data <- function(conn, service_type) {
  query <- sprintf(
    "SELECT *, Location AS location_code
    FROM [ANALYSIS_DB].[dbo].[service_utilization_%s]
    WHERE (service_date <= '2024-03-31')",
    service_type
  )
  dbGetQuery(conn, query)
}

included_locations <- c('LOC001', 'LOC002', 'LOC003', 'LOC004', 'LOC005', 
                       'LOC006', 'LOC007', 'LOC008', 'LOC009', 'LOC010', 
                       'LOC011', 'LOC012', 'LOC013', 'LOC014', 'LOC015', 'LOC016')

clean_data <- function(data, location_col) {
  data %>%
    mutate(
      socioeconomic_index = if_else(is.na(socioeconomic_index), "Unknown", as.character(socioeconomic_index)),
      gender = if_else(is.na(sex), "Unknown", sex),
      demographic_group = if_else(
        is.na(demographic_group) | demographic_group == "Unknown",
        "Unknown Group",
        demographic_group
      ),
      age_group = case_when(
        age < 25 ~ "16-24",
        age >= 25 & age < 35 ~ "25-34",
        age >= 35 & age < 50 ~ "35-49",
        age >= 50 & age < 65 ~ "50-64",
        age >= 65 & age < 80 ~ "65-79",
        age >= 80 ~ "80 and over",
        TRUE ~ NA_character_
      ),
      Location = as.character(!!sym(location_col))
    ) %>%
    filter(
      socioeconomic_index != "Unknown",
      gender != "Unknown",
      demographic_group != "Unknown Group",
      !is.na(age_group),
      !is.na(visit_count),
      Location %in% included_locations
    ) %>%
    mutate(
      gender = case_when(
        gender == "MALE" ~ "Men",
        gender == "FEMALE" ~ "Women",
        TRUE ~ gender
      ),
      age_group = str_replace_all(age_group, "-", " to "),
      age_group = str_replace_all(age_group, "80\\+", "80 and over"),
      age_group = str_replace_all(age_group, "(\\d+)(to)(\\d+)", "\\1 to \\3")
    )
}

service_a_data <- clean_data(load_service_data(db_conn, "type_a"), "Location")
service_b_data <- clean_data(load_service_data(db_conn, "type_b"), "Location")
service_c_data <- clean_data(load_service_data(db_conn, "type_c"), "Location")
```

```{r create-analysis-datasets}
create_analysis_dataset <- function(data, service_type) {
  data %>%
    mutate(
      service_test = case_when(
        lab_test == "Yes" & service_provided == "Yes" ~ 1,
        lab_test == "Yes" & (is.na(service_provided) | service_provided != "Yes") ~ 0,
        TRUE ~ NA_real_
      )
    ) %>%
    filter(!is.na(service_test)) %>%
    dplyr::select(
      participant_id, service_test,
      age_group, gender, demographic_group, socioeconomic_index, Location, visit_count
    )
}

service_a_analysis <- create_analysis_dataset(service_a_data, "type_a")
service_b_analysis <- create_analysis_dataset(service_b_data, "type_b")
service_c_analysis <- create_analysis_dataset(service_c_data, "type_c")

write.csv(service_a_analysis, "service_a_analysis.csv", row.names = FALSE)
write.csv(service_b_analysis, "service_b_analysis.csv", row.names = FALSE)
write.csv(service_c_analysis, "service_c_analysis.csv", row.names = FALSE)
```

```{r regression-setup}
reorder_factors <- function(data) {
  data %>%
    mutate(
      age_group = fct_relevel(as.factor(age_group), "35 to 49"),
      gender = fct_relevel(as.factor(gender), "Women"),
      demographic_group = fct_relevel(as.factor(demographic_group), "Group A"),
      Location = fct_relevel(as.factor(Location), "LOC001"),
      socioeconomic_index = fct_relevel(as.factor(socioeconomic_index), "1"),
      visit_count = fct_relevel(as.factor(visit_count), "1")
    )
}

reference_groups_uni <- data.frame(
  term = c("age_group 35 to 49 (reference)",
           "gender Women (reference)",
           "demographic_group Group A (reference)",
           "socioeconomic_index 1 (reference)",
           "Location LOC001 (reference)",
           "visit_count 1 (reference)")
)

calculate_odds_ratio <- function(sub_table) {
  or <- (sub_table[1, 1] * sub_table[2, 2]) / (sub_table[1, 2] * sub_table[2, 1])
  return(or)
}

univariable_analysis <- function(df, outcome, reference_groups) {
  predictor_vars <- c("age_group", "gender", "demographic_group", 
                     "socioeconomic_index", "Location", "visit_count")
  
  results <- lapply(predictor_vars, function(var) {
    formula <- as.formula(paste(outcome, "~", var))
    model <- glm(formula, data = df, family = binomial)
    
    coef_summary <- summary(model)$coefficients
    odds_ratios <- exp(coef_summary[, "Estimate"])
    p_values <- coef_summary[, "Pr(>|z|)"]
    
    odds_ratios[1] <- 1
    
    chi_sq <- anova(model, test = "Chisq")
    chi_sq_value <- chi_sq[2, "Deviance"]
    p_value <- chi_sq[2, "Pr(>Chi)"]
    
    table <- table(df[[var]], df[[outcome]])
    percentages <- prop.table(table, margin = 1) * 100
    combined <- cbind(table, round(percentages, 1))
    colnames(combined) <- c("No", "Yes", "No_%", "Yes_%")
    
    reference <- reference_groups_uni$term[grepl(var, reference_groups_uni$term)]
    
    detailed_results <- data.frame(
      Variable = var,
      Subgroup = rownames(combined),
      No = combined[,"No"],
      Yes = combined[,"Yes"],
      No_Percent = combined[,"No_%"],
      Yes_Percent = combined[,"Yes_%"],
      Chi2 = chi_sq_value,
      p_value_univariable = p_value,
      Odds_Ratio = odds_ratios,
      Reference_Group = reference
    ) %>%
      filter(Subgroup != "(Intercept)")
    
    return(detailed_results)
  })
  
  do.call(rbind, results)
}

logistic_regression <- function(df, outcome) {
  model <- glm(as.formula(paste(outcome, 
    "~ age_group + gender + demographic_group + socioeconomic_index + Location + visit_count")),
    data = df, family = binomial)
  
  coef_summary <- summary(model)$coefficients
  ci <- confint.default(model)
  
  tidy_model <- data.frame(
    term = rownames(coef_summary),
    OR = exp(coef_summary[, "Estimate"]),
    CI_lower = exp(ci[, 1]),
    CI_upper = exp(ci[, 2]),
    p.value_multivariable = coef_summary[, "Pr(>|z|)"]
  ) %>%
    filter(term != "(Intercept)")
  
  return(tidy_model)
}
```



```{r multilevel-regression-analysis}
# Basic Model Analysis
# Load dataset
df_main <- read_csv("service_analysis.csv")

# Ensure factors are properly set
df_main$location <- as.factor(df_main$location)
df_main$visit_count <- as.numeric(as.character(df_main$visit_count))
df_main$service_test <- as.numeric(as.character(df_main$service_test))

# Create visit categories
df_main <- df_main %>%
  mutate(
    visit_count_cat = case_when(
      visit_count == 1 ~ "1",
      visit_count == 2 ~ "2",
      visit_count >= 3 & visit_count <= 5 ~ "3-5",
      visit_count >= 6 & visit_count <= 10 ~ "6-10",
      visit_count >= 11 ~ "11+"
    )
  ) %>%
  mutate(visit_count_cat = factor(visit_count_cat, 
                                 levels = c("1", "2", "3-5", "6-10", "11+")))

# Set reference levels
df_main <- df_main %>%
  mutate(
    age_group = fct_relevel(as.factor(age_group), "35 to 49"),
    gender = fct_relevel(as.factor(gender), "Primary"),
    demographic_group = fct_relevel(as.factor(demographic_group), "Group A"),
    socioeconomic_index = fct_relevel(as.factor(socioeconomic_index), "1"),
    visit_count_cat = fct_relevel(as.factor(visit_count_cat), "1")
  )

# Fit base mixed model
mixed_model <- glmer(
  service_test ~ age_group + gender + demographic_group + 
    socioeconomic_index + visit_count_cat + (1 | location),
  data = df_main,
  family = binomial,
  nAGQ = 0
)

# Process results
tidy_mixed_model <- broom.mixed::tidy(
  mixed_model,
  conf.int = TRUE,
  conf.method = "Wald",
  effects = "fixed"
) %>%
  mutate(
    OR = exp(estimate),
    OR_lower = exp(conf.low),
    OR_upper = exp(conf.high)
  )

# Gender-Specific Analysis
# Fit model with random slopes for gender
mixed_model_gender <- glmer(
  service_test ~ age_group + gender + demographic_group + 
    socioeconomic_index + visit_count_cat + (1 + gender | location),
  data = df_main,
  family = binomial,
  nAGQ = 0,
  control = glmerControl(optimizer = "bobyqa", 
                        optCtrl = list(maxfun = 1e5))
)

# Extract gender-specific results
gender_results <- broom.mixed::tidy(
  mixed_model_gender,
  conf.int = TRUE,
  conf.method = "Wald",
  effects = "fixed"
) %>%
  mutate(
    OR = exp(estimate),
    OR_lower = exp(conf.low),
    OR_upper = exp(conf.high)
  )

# Age Group Analysis
# Create age group indicators
df_main <- df_main %>%
  mutate(
    age_16_24 = as.numeric(age_group == "16 to 24"),
    age_25_34 = as.numeric(age_group == "25 to 34"),
    age_50_64 = as.numeric(age_group == "50 to 64"),
    age_65_79 = as.numeric(age_group == "65 to 79"),
    age_80plus = as.numeric(age_group == "80 and over")
  )

# Fit age group model
mixed_model_age <- glmer(
  service_test ~ age_group + gender + demographic_group + 
    socioeconomic_index + visit_count_cat + 
    (1 + age_16_24 + age_25_34 + age_50_64 + 
     age_65_79 + age_80plus || location),
  data = df_main,
  family = binomial,
  nAGQ = 0,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 1e5),
    calc.derivs = FALSE
  )
)

# Process age group results
age_random_effects <- ranef(mixed_model_age)$location
age_variations <- data.frame(
  location = rownames(age_random_effects),
  Intercept = age_random_effects[,1],
  Effect_16_24 = age_random_effects[,2],
  Effect_25_34 = age_random_effects[,3],
  Effect_50_64 = age_random_effects[,4],
  Effect_65_79 = age_random_effects[,5],
  Effect_80plus = age_random_effects[,6]
)

# Socioeconomic Analysis
# Create index indicators
df_main <- df_main %>%
  mutate(
    index_2 = as.numeric(socioeconomic_index == "2"),
    index_3 = as.numeric(socioeconomic_index == "3"),
    index_4 = as.numeric(socioeconomic_index == "4"),
    index_5 = as.numeric(socioeconomic_index == "5")
  )

# Fit socioeconomic model
mixed_model_socioeconomic <- glmer(
  service_test ~ socioeconomic_index + gender + age_group + 
    demographic_group + visit_count_cat + 
    (1 + index_2 + index_3 + index_4 + index_5 || location),
  data = df_main,
  family = binomial,
  nAGQ = 0,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 1e5),
    calc.derivs = FALSE
  )
)

# Demographic Group Analysis
# Create demographic indicators
df_main <- df_main %>%
  mutate(
    group_b = as.numeric(demographic_group == "Group B"),
    group_c = as.numeric(demographic_group == "Group C"),
    group_d = as.numeric(demographic_group == "Group D"),
    group_e = as.numeric(demographic_group == "Group E"),
    group_f = as.numeric(demographic_group == "Group F"),
    group_g = as.numeric(demographic_group == "Group G"),
    group_h = as.numeric(demographic_group == "Group H"),
    group_i = as.numeric(demographic_group == "Group I")
  )

# Fit demographic model
mixed_model_demographic <- glmer(
  service_test ~ demographic_group + gender + age_group + 
    socioeconomic_index + visit_count_cat + 
    (1 + group_b + group_c + group_d + group_e + 
     group_f + group_g + group_h + group_i || location),
  data = df_main,
  family = binomial,
  nAGQ = 0,
  control = glmerControl(
    optimizer = "bobyqa",
    optCtrl = list(maxfun = 1e5),
    calc.derivs = FALSE
  )
)

# Process demographic results
demographic_random_effects <- ranef(mixed_model_demographic)$location
demographic_variations <- data.frame(
  location = rownames(demographic_random_effects),
  Intercept = demographic_random_effects[,1],
  Effect_Group_B = demographic_random_effects[,2],
  Effect_Group_C = demographic_random_effects[,3],
  Effect_Group_D = demographic_random_effects[,4],
  Effect_Group_E = demographic_random_effects[,5],
  Effect_Group_F = demographic_random_effects[,6],
  Effect_Group_G = demographic_random_effects[,7],
  Effect_Group_H = demographic_random_effects[,8],
  Effect_Group_I = demographic_random_effects[,9]
)

# Generate Visualizations
# Forest plot for main effects
forest_plot_main <- ggplot(
  tidy_mixed_model %>% 
    filter(term != "(Intercept)" & !str_detect(term, "^visit_count_cat")),
  aes(x = OR, y = term)
) +
  geom_point() +
  geom_errorbarh(aes(xmin = OR_lower, xmax = OR_upper), height = 0.2) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  xlab("Odds Ratio") +
  ylab("Predictor Variables") +
  ggtitle("Forest Plot of Mixed Effects Results") +
  theme_pubr() +
  theme(axis.text.y = element_text(hjust = 1, size = 10))

# Save results
write.csv(tidy_mixed_model, "mixed_model_results.csv", row.names = FALSE)
write.csv(demographic_variations, "demographic_variations.csv", row.names = FALSE)
write.csv(age_variations, "age_variations.csv", row.names = FALSE)

ggsave("forest_plot_main.png", plot = forest_plot_main, width = 10, height = 8)

# Plot random effects
random_effects_plot <- dotplot(ranef(mixed_model, condVar = TRUE), 
                             main = "Random Effects by Location")
ggsave("random_effects.png", plot = random_effects_plot, width = 10, height = 6)

```



# Load required libraries
library(tidyverse)
library(ggplot2)
library(gridExtra)
library(ggpubr)
library(caret)
library(pROC)
library(stats)

#######################################
# Part 1: Analysis for Outcome A
#######################################

# Load and prepare the dataset
df_outcome_a <- read_csv("outcome_a_data.csv")
df_outcome_a <- reorder_factors(df_outcome_a)

# Univariable analysis
uni_results_a <- univariable_analysis(df_outcome_a, "outcome_a", reference_groups_uni)
uni_results_a <- uni_results_a %>%
  mutate(across(where(is.numeric), round, 3))

# Save univariable results table
table_grob_a_univariable <- tableGrob(uni_results_a, rows = NULL)
ggsave("results_table_outcome_a_univariable.png", 
       plot = table_grob_a_univariable, 
       width = 15, 
       height = 14)

# Perform logistic regression
logit_results_a <- logistic_regression(df_outcome_a, "outcome_a")

# Define reference groups
reference_groups <- data.frame(
  term = c("age_category_3 (reference)", 
           "sex_category_1 (reference)", 
           "demographic_group_1 (reference)", 
           "social_index_1 (reference)", 
           "location_1 (reference)",
           "visit_count_1 (reference)"),
  OR = NA, 
  CI_lower = NA, 
  CI_upper = NA, 
  p.value_multivariable = NA
)

# Combine results with reference groups
combined_results_a <- bind_rows(reference_groups, logit_results_a) %>%
  mutate(term = make.unique(as.character(term))) %>%
  distinct(term, .keep_all = TRUE) %>%
  mutate(across(where(is.numeric), round, 3))

# Print results
print("Results for Outcome A")
print(combined_results_a)

# Save multivariable results table
table_grob_a <- tableGrob(combined_results_a, rows = NULL)
ggsave("results_table_outcome_a_multivariable.png", 
       plot = table_grob_a, 
       width = 10, 
       height = 16)

# Prepare and create forest plot
forest_plot_data <- combined_results_a %>%
  dplyr::filter(!str_detect(term, "^location")) %>%
  dplyr::filter(!str_detect(term, "^visit_count")) %>%
  dplyr::filter(!str_detect(term, "^Intercept|NA")) %>%
  mutate(term = case_when(
    term == "social_index_1" ~ "Social Index Level 1",
    term == "social_index_2" ~ "Social Index Level 2",
    term == "social_index_3" ~ "Social Index Level 3",
    term == "social_index_4" ~ "Social Index Level 4",
    term == "social_index_5" ~ "Social Index Level 5",
    term == "sex_category_1" ~ "Sex Category 1 (reference)",
    term == "sex_category_2" ~ "Sex Category 2",
    term == "demographic_group_1" ~ "Demographic Group 1 (reference)",
    term == "demographic_group_2" ~ "Demographic Group 2",
    term == "demographic_group_3" ~ "Demographic Group 3",
    term == "demographic_group_4" ~ "Demographic Group 4",
    term == "demographic_group_5" ~ "Demographic Group 5",
    term == "demographic_group_6" ~ "Demographic Group 6",
    term == "demographic_group_7" ~ "Demographic Group 7",
    term == "demographic_group_8" ~ "Demographic Group 8",
    term == "demographic_group_9" ~ "Demographic Group 9",
    term == "age_category_1" ~ "Age Category 1",
    term == "age_category_2" ~ "Age Category 2",
    term == "age_category_3" ~ "Age Category 3 (reference)",
    term == "age_category_4" ~ "Age Category 4",
    term == "age_category_5" ~ "Age Category 5",
    term == "age_category_6" ~ "Age Category 6",
    TRUE ~ term
  ))

forest_plot_a <- ggplot(forest_plot_data, aes(x = OR, y = term)) +
  geom_point() +
  geom_errorbarh(aes(xmin = CI_lower, xmax = CI_upper)) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  xlab("Odds Ratio") +
  ylab("Predictor Variables") +
  ggtitle("Forest Plot of Outcome A Logistic Regression Results") +
  theme_pubr() +
  theme(axis.text.y = element_text(hjust = 1, size = 10))

ggsave("forest_plot_outcome_a.png", 
       plot = forest_plot_a, 
       width = 10, 
       height = 8)

#######################################
# Part 2: Analysis for Outcome B
#######################################

# Load and prepare the dataset
df_outcome_b <- read_csv("outcome_b_data.csv")
df_outcome_b <- reorder_factors(df_outcome_b)

# Univariable analysis
uni_results_b <- univariable_analysis(df_outcome_b, "outcome_b", reference_groups_uni)
uni_results_b <- uni_results_b %>%
  mutate(across(where(is.numeric), round, 3))

# Save univariable results table
table_grob_b_univariable <- tableGrob(uni_results_b, rows = NULL)
ggsave("results_table_outcome_b_univariable.png", 
       plot = table_grob_b_univariable, 
       width = 15, 
       height = 14)

# Perform logistic regression and combine results
logit_results_b <- logistic_regression(df_outcome_b, "outcome_b")
combined_results_b <- bind_rows(reference_groups, logit_results_b) %>%
  mutate(term = make.unique(as.character(term))) %>%
  distinct(term, .keep_all = TRUE) %>%
  mutate(
    CI_lower = as.numeric(CI_lower),
    CI_upper = as.numeric(CI_upper)
  ) %>%
  mutate(across(where(is.numeric) & !matches("CI_upper"), round, 3),
         CI_upper = round(CI_upper, 3)
  )

# Print results
print("Results for Outcome B")
print(combined_results_b)

# Save multivariable results table
table_grob_b <- tableGrob(combined_results_b, rows = NULL)
ggsave("results_table_outcome_b_multivariable.png", 
       plot = table_grob_b, 
       width = 10, 
       height = 16)

#######################################
# Part 3: Nested Analysis
#######################################

# Load dataset
df_outcome_c <- read_csv("outcome_c_data.csv")

# Fit base and extended models
model_base <- glm(outcome_c ~ age_category + sex_category + demographic_group + 
                   social_index + location, 
                 data = df_outcome_c, 
                 family = binomial)

model_extended <- glm(outcome_c ~ age_category + sex_category + demographic_group + 
                       social_index + location + visit_count_grouped, 
                     data = df_outcome_c, 
                     family = binomial)

# Perform likelihood ratio test
lrt_result <- anova(model_base, model_extended, test = "LRT")
print("Likelihood Ratio Test results:")
print(lrt_result)

# Compare AIC
aic_comparison <- data.frame(
  Model = c("Base model", "Extended model"),
  AIC = c(AIC(model_base), AIC(model_extended))
)
print("\nAIC Comparison:")
print(aic_comparison)

# Calculate and compare McFadden's R-squared
mcfadden_r2 <- function(model) {
  1 - (model$deviance / model$null.deviance)
}

r2_comparison <- data.frame(
  Model = c("Base model", "Extended model"),
  McFadden_R2 = c(mcfadden_r2(model_base), mcfadden_r2(model_extended))
)
print("\nMcFadden's R-squared Comparison:")
print(r2_comparison)

# Compare number of parameters
params_comparison <- data.frame(
  Model = c("Base model", "Extended model"),
  Number_of_Parameters = c(length(coef(model_base)), 
                          length(coef(model_extended)))
)
print("\nNumber of Parameters Comparison:")
print(params_comparison)

# Perform cross-validation
set.seed(123)
cv_folds <- createFolds(df_outcome_c$outcome_c, k = 10)

cv_results <- lapply(cv_folds, function(test_indices) {
  train_data <- df_outcome_c[-test_indices, ]
  test_data <- df_outcome_c[test_indices, ]
  
  # Train and test base model
  model_base_cv <- glm(outcome_c ~ age_category + sex_category + 
                        demographic_group + social_index + location, 
                      data = train_data, 
                      family = binomial)
  base_pred <- predict(model_base_cv, newdata = test_data, type = "response")
  base_auc <- auc(test_data$outcome_c, base_pred)
  
  # Train and test extended model
  model_extended_cv <- glm(outcome_c ~ age_category + sex_category + 
                            demographic_group + social_index + location + 
                            visit_count_grouped, 
                         data = train_data, 
                         family = binomial)
  extended_pred <- predict(model_extended_cv, newdata = test_data, type = "response")
  extended_auc <- auc(test_data$outcome_c, extended_pred)
  
  return(c(base_auc, extended_auc))
})

cv_results_df <- data.frame(do.call(rbind, cv_results))
colnames(cv_results_df) <- c("Base Model", "Extended Model")

print("\nCross-validation Results (AUC):")
print(cv_results_df)
print("\nMean AUC:")
print(colMeans(cv_results_df))

#######################################
# Part 4: Stratified Analysis
#######################################

# Define automated locations
automated_locations <- c('LOC001', 'LOC002', 'LOC003', 'LOC004', 'LOC005', 
                        'LOC006', 'LOC007', 'LOC008', 'LOC009')

# Create automation status variable and split data
df_outcome_a <- df_outcome_a %>%
  mutate(automation_status = ifelse(location %in% automated_locations, 
                                  "Automated", "Manual"))

df_automated <- df_outcome_a %>% filter(automation_status == "Automated")
df_manual <- df_outcome_a %>% filter(automation_status == "Manual")

# Fit stratified models
model_automated <- glm(outcome_a ~ age_category + sex_category + demographic_group + 
                        social_index + visit_count_grouped + location, 
                      data = df_automated, 
                      family = binomial)

model_manual <- glm(outcome_a ~ age_category + sex_category + demographic_group + 
                     social_index + visit_count_grouped + location, 
                   data = df_manual, 
                   family = binomial)

# Print model summaries
print("Model Summary for Automated Locations:")
print(summary(model_automated))
print("\nModel Summary for Manual Locations:")
print(summary(model_manual))

# Perform cross-validation for stratified models
set.seed(123)
cv_folds_auto <- createFolds(df_automated$outcome_a, k = 10)
cv_folds_manual <- createFolds(df_manual$outcome_a, k = 10)

# Cross-validation for automated locations
cv_results_auto <- lapply(cv_folds_auto, function(test_indices) {
  train_data <- df_automated[-test_indices, ]
  test_data <- df_automated[test_indices, ]
  
  model_auto_cv <- glm(outcome_a ~ age_category + sex_category + demographic_group + 
                        social_index + visit_count_grouped + location, 
                      data = train_data, 
                      family = binomial)
  auto_pred <- predict(model_auto_cv, newdata = test_data, type = "response")
  auto_auc <- auc(test_data$outcome_a, auto_pred)
  
  return(auto_auc)
})

# Cross-validation for manual locations
cv_results_manual <- lapply(cv_folds_manual, function(test_indices) {
  train_data <- df_manual[-test_indices, ]
  test_data <- df_manual[test_indices, ]
  
  model_manual_cv <- glm(outcome_a ~ age_category + sex_category + demographic_group + 
                          social_index + visit_count_grouped + location, 
                        data = train_data, 
                        family = binomial)
  manual_pred <- predict(model_manual_cv, newdata = test_data, type = "response")
  manual_auc <- auc(test_data$outcome_a, manual_pred)
  
  return(manual_auc)
})

# Compile and print cross-validation results
cv_results_df_auto <- data.frame(AUC_Automated = unlist(cv_results_auto))
cv_results_df_manual <- data.frame(AUC_Manual = unlist(cv_results_manual))

print("\nCross-validation Results (AUC) for Automated Locations:")
print(cv_results_df_auto)
print("\nCross-validation Results (AUC) for Manual Locations:")
print(cv_results_df_manual)
print("\nMean AUC for Automated Locations:")
print(mean(cv_results_df_auto$AUC_Automated))
print("\nMean AUC for Manual Locations:")
print(mean(cv_results_df_manual$AUC_Manual))


```


