---
title: "data_quality_per_site"
author: "Jessica Edney"
date: "2024-03-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
knitr::opts_knit$set(root.dir = 'Y:/NEW NCSP/RESEARCH AND DEVELOPMENT/J Edney/ED opt out/ed-monitoring/')
```

# Overall comparison

text text
```{r packages, echo=FALSE}
# load packages
pacman::p_load(
    pacman,
    tidyverse,
    lubridate,
    readxl,
    writexl,
    reshape2,
    scales,
    flextable,
    officer
)
```

```{r import, echo=FALSE}
# run data comparison script (make take a while)
source("data_comparison.R")
source("theme_ukhsa.R")
source("ukhsa_colours.R")

pal_hc <- ukhsa_colours("high_contrast")

options(scipen = 999)
```

```{r data prep, echo=FALSE}
overall <- sentinel_nhse_compared %>%
group_by(event) %>%
mutate(NHSE = sum(nhse_total),
ECDS = sum(ecds_total)) %>%
select(event, NHSE, ECDS) %>%
distinct() %>% 
mutate(`Difference (%)` = round(((ECDS-NHSE)/NHSE) * 100, 1),
       `Difference (absolute)` = ECDS-NHSE) %>%
mutate(`Difference (%)` = `Difference (%)`^2,
       `Difference (absolute)` = `Difference (absolute)`^2) %>%
mutate(`Difference (%)` = sqrt(`Difference (%)`),
       `Difference (absolute)` = sqrt(`Difference (absolute)`)) %>%
ungroup()

overall$Event <- str_wrap(overall$event, width = 10)
```

### Fig.1a: Difference (%) in number of events compared to NHSE data by event type
```{r difference, echo=FALSE, fig.width=7, fig.height=5}
ggplot(overall, aes(x = reorder(Event, -`Difference (%)`), y = `Difference (%)`, fill = `Difference (%)`)) +
geom_col(aes(fill = `Difference (%)`)) +
scale_y_continuous(labels = label_comma(), expand = c(0, 0)) +
scale_fill_continuous(trans = 'reverse') +
labs(x = "Event")+
theme_ukhsa() +
theme(legend.position = "none",
  text = element_text(size=12),
  plot.subtitle = element_text(vjust = 5))
```

### Fig.1b: Difference (absolute) in number of events compared to NHSE data by event type
```{r difference2, echo=FALSE, fig.width=7, fig.height=5}
ggplot(overall, aes(x = reorder(Event, -`Difference (absolute)`), y = `Difference (absolute)`, fill = `Difference (absolute)`)) +
geom_col(aes(fill = `Difference (absolute)`)) +
scale_y_continuous(labels = label_comma(), expand = c(0, 0)) +
scale_fill_continuous(trans = 'reverse') +
labs(x = "Event")+
theme_ukhsa() +
theme(legend.position = "none",
  text = element_text(size=12),
  plot.subtitle = element_text(vjust = 5))
```

# Attendances 

```{r attendances, echo=FALSE}
attendances <- sentinel_nhse_compared %>%
filter(event == "Attendance") %>%
select(-event) %>%
rename(Site = site_code,
NHSE = nhse_total,
ECDS = ecds_total)

attendances$perc_diff <- as.numeric(attendances$perc_diff)
                                    
attendances <- attendances %>%
  mutate(perc_diff = round(perc_diff, 1)) %>%
  rename(`Difference (%)` = perc_diff,
         `Difference (absolute)` = abs_diff)
       
# define values
mean_nhse <- round(mean(attendances$NHSE), 0)
total_nhse <- round(sum(attendances$NHSE), 0)
mean_ecds <- round(mean(attendances$ECDS), 0)
total_ecds <- round(sum(attendances$ECDS), 0)
mean_diff <- round(mean(attendances$perc_diff), 0)
max_diff <- attendances[which.max(attendances$perc_diff),]
min_diff <- attendances[which.min(attendances$perc_diff),]
```

### Fig.2a: Difference (%) in number of ECDS ED attendances compared to NHSE data by site 
```{r attendances_graph, echo=FALSE, fig.width=7, fig.height=5}
ggplot(attendances, aes(x = reorder(Site, -`Difference (%)`), y = `Difference (%)`)) +
geom_col(aes(fill = `Difference (%)`), position = "dodge2") +
scale_y_continuous(labels = label_comma(), expand = c(0, 0)) +
scale_fill_continuous(trans = 'reverse') +
labs(x = "Site") +
theme_ukhsa() +
theme(legend.position = "none",
  text = element_text(size=12),
  legend.text = element_text(size=12),
  axis.text.x = element_text(angle = 90),
  plot.subtitle = element_text(vjust = 5))
```

### Fig.2b: Difference (absolute) in number of ECDS ED attendances compared to NHSE data by site 
```{r attendances_graph2, echo=FALSE, fig.width=7, fig.height=5}
ggplot(attendances, aes(x = reorder(Site, -`Difference (absolute)`), y = `Difference (absolute)`)) +
geom_col(aes(fill = `Difference (absolute)`), position = "dodge2") +
scale_y_continuous(labels = label_comma(), expand = c(0, 0)) +
scale_fill_continuous(trans = 'reverse') +
labs(x = "Site") +
theme_ukhsa() +
theme(legend.position = "none",
  text = element_text(size=12),
  legend.text = element_text(size=12),
  axis.text.x = element_text(angle = 90),
  plot.subtitle = element_text(vjust = 5))
```

### Table 1: Numbers of ED attendances in NHSE and ECDS data by site with percentage difference (ECDS compared to NHSE)
```{r attendances_table, echo=FALSE}
# green if less than 20%
# yellow if 21-30%
# red if over 30%

attendances_tbl <- attendances %>%
  arrange(`Difference (%)`) %>%
  flextable() %>%
  autofit() %>%
  bg(., i = ~`Difference (%)` < 20, part = "body", bg = "green") %>%
  bg(., i = ~`Difference (%)` >= 20 & `Difference (%)` < 30, part = "body", bg = "yellow") %>%
  bg(., i = ~`Difference (%)` >= 30, part = "body", bg = "red")
  
attendances_tbl
```

# Blood tests

```{r blood_tests, echo=FALSE}
bloodtests <- sentinel_nhse_compared %>%
filter(event == "Blood test") %>%
select(-event) %>%
rename(Site = site_code,
NHSE = nhse_total,
ECDS = ecds_total)

bloodtests$perc_diff <- as.numeric(bloodtests$perc_diff)
                                    
bloodtests <- bloodtests %>%
  mutate(perc_diff = round(perc_diff, 1)) %>%
  rename(`Difference (%)` = perc_diff,
         `Difference (absolute)` = abs_diff)
       
# define values
mean_nhse <- round(mean(bloodtests$NHSE), 0)
total_nhse <- round(sum(bloodtests$NHSE), 0)
mean_ecds <- round(mean(bloodtests$ECDS), 0)
total_ecds <- round(sum(bloodtests$ECDS), 0)
mean_diff <- round(mean(bloodtests$perc_diff), 0)
max_diff <- bloodtests[which.max(bloodtests$perc_diff),]
min_diff <- bloodtests[which.min(bloodtests$perc_diff),]
```

### Fig.3a: Difference (%) in number of ECDS ED attendances with blood tests compared to NHSE data
```{r bloodtests_graph, echo=FALSE, fig.width=7, fig.height=5}
ggplot(bloodtests, aes(x = reorder(Site, -`Difference (%)`), y = `Difference (%)`)) +
geom_col(aes(fill = `Difference (%)`), position = "dodge2") +
scale_y_continuous(labels = label_comma(), expand = c(0, 0)) +
scale_fill_continuous(trans = 'reverse') +
labs(x = "Site") +
theme_ukhsa() +
theme(legend.position = "none",
  text = element_text(size=12),
  legend.text = element_text(size=12),
  axis.text.x = element_text(angle = 90),
  plot.subtitle = element_text(vjust = 5))
```

### Fig.3b: Difference (absolute) in number of ECDS ED attendances with blood tests compared to NHSE data
```{r bloodtests_graph2, echo=FALSE, fig.width=7, fig.height=5}
ggplot(bloodtests, aes(x = reorder(Site, -`Difference (absolute)`), y = `Difference (absolute)`)) +
geom_col(aes(fill = `Difference (absolute)`), position = "dodge2") +
scale_y_continuous(labels = label_comma(), expand = c(0, 0)) +
scale_fill_continuous(trans = 'reverse') +
labs(x = "Site") +
theme_ukhsa() +
theme(legend.position = "none",
  text = element_text(size=12),
  legend.text = element_text(size=12),
  axis.text.x = element_text(angle = 90),
  plot.subtitle = element_text(vjust = 5))
```

### Table 2: Numbers of ED attendances with blood tests in NHSE and ECDS data by site with percentage difference (ECDS compared to NHSE)
```{r bloodtests_table, echo=FALSE}
# green if less than 20%
# yellow if 21-30%
# red if over 30%
bloodtests_tbl <- bloodtests %>%
  arrange(`Difference (%)`) %>%
  flextable() %>%
  autofit() %>%
  bg(., i = ~`Difference (%)` < 20, part = "body", bg = "green") %>%
  bg(., i = ~`Difference (%)` >= 20 & `Difference (%)` < 30, part = "body", bg = "yellow") %>%
  bg(., i = ~`Difference (%)` >= 30, part = "body", bg = "red")
  
bloodtests_tbl
```

# HBV tests

```{r hbv_tests, echo=FALSE}
hbvtests <- sentinel_nhse_compared %>%
filter(event == "HBV test") %>%
select(-event) %>%
rename(Site = site_code,
NHSE = nhse_total,
ECDS = ecds_total)

hbvtests$perc_diff <- as.numeric(hbvtests$perc_diff)
                                    
hbvtests <- hbvtests %>%
  mutate(perc_diff = round(perc_diff, 1)) %>%
  rename(`Difference (%)` = perc_diff,
         `Difference (absolute)` = abs_diff)
       
# define values
mean_nhse <- round(mean(hbvtests$NHSE), 0)
total_nhse <- round(sum(hbvtests$NHSE), 0)
mean_ecds <- round(mean(hbvtests$ECDS), 0)
total_ecds <- round(sum(hbvtests$ECDS), 0)
mean_diff <- round(mean(hbvtests$perc_diff), 0)
max_diff <- hbvtests[which.max(hbvtests$perc_diff),]
min_diff <- hbvtests[which.min(hbvtests$perc_diff),]
```

### Fig.4a: Difference (%) in number of ECDS ED attendances with HBV tests compared to NHSE data
```{r hbvtests_graph, echo=FALSE, fig.width=7, fig.height=5}
ggplot(hbvtests, aes(x = reorder(Site, -`Difference (%)`), y = `Difference (%)`)) +
geom_col(aes(fill = `Difference (%)`), position = "dodge2") +
scale_y_continuous(labels = label_comma(), expand = c(0, 0)) +
scale_fill_continuous(trans = 'reverse') +
labs(x = "Site") +
theme_ukhsa() +
theme(legend.position = "none",
  text = element_text(size=12),
  legend.text = element_text(size=12),
  axis.text.x = element_text(angle = 90),
  plot.subtitle = element_text(vjust = 5))
```

### Fig.4b: Difference (absolute) in number of ECDS ED attendances with HBV tests compared to NHSE data
```{r hbvtests_graph2, echo=FALSE, fig.width=7, fig.height=5}
ggplot(hbvtests, aes(x = reorder(Site, -`Difference (absolute)`), y = `Difference (absolute)`)) +
geom_col(aes(fill = `Difference (absolute)`), position = "dodge2") +
scale_y_continuous(labels = label_comma(), expand = c(0, 0)) +
scale_fill_continuous(trans = 'reverse') +
labs(x = "Site") +
theme_ukhsa() +
theme(legend.position = "none",
  text = element_text(size=12),
  legend.text = element_text(size=12),
  axis.text.x = element_text(angle = 90),
  plot.subtitle = element_text(vjust = 5))
```

### Table 3: Numbers of ED attendances with HBV tests in NHSE and ECDS data by site with percentage difference (ECDS compared to NHSE)
```{r hbvtests_table, echo=FALSE}
# green if less than 20%
# yellow if 21-30%
# red if over 30%
hbvtests_tbl <- hbvtests %>%
  arrange(`Difference (%)`) %>%
  flextable() %>%
  autofit() %>%
  bg(., i = ~`Difference (%)` < 20, part = "body", bg = "green") %>%
  bg(., i = ~`Difference (%)` >= 20 & `Difference (%)` < 30, part = "body", bg = "yellow") %>%
  bg(., i = ~`Difference (%)` >= 30 | NHSE == 0, part = "body", bg = "red")
  
hbvtests_tbl
```

# HCV tests

```{r hcv_tests, echo=FALSE}
hcvtests <- sentinel_nhse_compared %>%
filter(event == "HCV test") %>%
select(-event) %>%
rename(Site = site_code,
NHSE = nhse_total,
ECDS = ecds_total)

hcvtests$perc_diff <- as.numeric(hcvtests$perc_diff)
                                    
hcvtests <- hcvtests %>%
  mutate(perc_diff = round(perc_diff, 1)) %>%
  rename(`Difference (%)` = perc_diff,
         `Difference (absolute)` = abs_diff)
       
# define values
mean_nhse <- round(mean(hcvtests$NHSE), 0)
total_nhse <- round(sum(hcvtests$NHSE), 0)
mean_ecds <- round(mean(hcvtests$ECDS), 0)
total_ecds <- round(sum(hcvtests$ECDS), 0)
mean_diff <- round(mean(hcvtests$perc_diff), 0)
max_diff <- hcvtests[which.max(hcvtests$perc_diff),]
min_diff <- hcvtests[which.min(hcvtests$perc_diff),]
```

### Fig.5a: Difference (%) in number of ECDS ED HCV tests compared to NHSE data
```{r hcvtests_graph, echo=FALSE, fig.width=7, fig.height=5}
ggplot(hcvtests, aes(x = reorder(Site, -`Difference (%)`), y = `Difference (%)`)) +
geom_col(aes(fill = `Difference (%)`), position = "dodge2") +
scale_y_continuous(labels = label_comma(), expand = c(0, 0)) +
scale_fill_continuous(trans = 'reverse') +
labs(x = "Site") +
theme_ukhsa() +
theme(legend.position = "none",
  text = element_text(size=12),
  legend.text = element_text(size=12),
  axis.text.x = element_text(angle = 90),
  plot.subtitle = element_text(vjust = 5))
```

### Fig.5b: Difference (absolute) in number of ECDS ED HCV tests compared to NHSE data
```{r hcvtests_graph2, echo=FALSE, fig.width=7, fig.height=5}
ggplot(hcvtests, aes(x = reorder(Site, -`Difference (absolute)`), y = `Difference (absolute)`)) +
geom_col(aes(fill = `Difference (absolute)`), position = "dodge2") +
scale_y_continuous(labels = label_comma(), expand = c(0, 0)) +
scale_fill_continuous(trans = 'reverse') +
labs(x = "Site") +
theme_ukhsa() +
theme(legend.position = "none",
  text = element_text(size=12),
  legend.text = element_text(size=12),
  axis.text.x = element_text(angle = 90),
  plot.subtitle = element_text(vjust = 5))
```

### Table 4: Numbers of ED attendances with HCV tests in NHSE and ECDS data by site with percentage difference (ECDS compared to NHSE)
```{r hcvtests_table, echo=FALSE}
# green if less than 20%
# yellow if 21-30%
# red if over 30%
hcvtests_tbl <- hcvtests %>%
  arrange(`Difference (%)`) %>%
  flextable() %>%
  autofit() %>%
  bg(., i = ~`Difference (%)` < 20, part = "body", bg = "green") %>%
  bg(., i = ~`Difference (%)` >= 20 & `Difference (%)` < 30, part = "body", bg = "yellow") %>%
  bg(., i = ~`Difference (%)` >= 30 | NHSE == 0, part = "body", bg = "red")
  
hcvtests_tbl
```

# HIV tests

```{r hiv_tests, echo=FALSE}
hivtests <- sentinel_nhse_compared %>%
filter(event == "HIV test") %>%
select(-event) %>%
rename(Site = site_code,
NHSE = nhse_total,
ECDS = ecds_total)

hivtests$perc_diff <- as.numeric(hivtests$perc_diff)
                                    
hivtests <- hivtests %>%
  mutate(perc_diff = round(perc_diff, 1)) %>%
  rename(`Difference (%)` = perc_diff,
         `Difference (absolute)` = abs_diff)
       
# define values
mean_nhse <- round(mean(hivtests$NHSE), 0)
total_nhse <- round(sum(hivtests$NHSE), 0)
mean_ecds <- round(mean(hivtests$ECDS), 0)
total_ecds <- round(sum(hivtests$ECDS), 0)
mean_diff <- round(mean(hivtests$perc_diff), 0)
max_diff <- hivtests[which.max(hivtests$perc_diff),]
min_diff <- hivtests[which.min(hivtests$perc_diff),]
```

### Fig.6a: Difference (%) in number of ECDS ED HIV tests compared to NHSE data
```{r hivtests_graph, echo=FALSE, fig.width=7, fig.height=5}
ggplot(hivtests, aes(x = reorder(Site, -`Difference (%)`), y = `Difference (%)`)) +
geom_col(aes(fill = `Difference (%)`), position = "dodge2") +
scale_y_continuous(labels = label_comma(), expand = c(0, 0)) +
scale_fill_continuous(trans = 'reverse') +
labs(x = "Site") +
theme_ukhsa() +
theme(legend.position = "none",
  text = element_text(size=12),
  legend.text = element_text(size=12),
  axis.text.x = element_text(angle = 90),
  plot.subtitle = element_text(vjust = 5))
```

### Fig.6b: Difference (absolute) in number of ECDS ED HIV tests compared to NHSE data
```{r hivtests_graph2, echo=FALSE, fig.width=7, fig.height=5}
ggplot(hivtests, aes(x = reorder(Site, -`Difference (absolute)`), y = `Difference (absolute)`)) +
geom_col(aes(fill = `Difference (absolute)`), position = "dodge2") +
scale_y_continuous(labels = label_comma(), expand = c(0, 0)) +
scale_fill_continuous(trans = 'reverse') +
labs(x = "Site") +
theme_ukhsa() +
theme(legend.position = "none",
  text = element_text(size=12),
  legend.text = element_text(size=12),
  axis.text.x = element_text(angle = 90),
  plot.subtitle = element_text(vjust = 5))
```

### Table 5: Numbers of ED attendances with HIV tests in NHSE and ECDS data by site with percentage difference (ECDS compared to NHSE)
```{r hivtests_table, echo=FALSE}
# green if less than 20%
# yellow if 21-30%
# red if over 30%
hivtests_tbl <- hivtests %>%
  arrange(`Difference (%)`) %>%
  flextable() %>%
  autofit() %>%
  bg(., i = ~`Difference (%)` < 20, part = "body", bg = "green") %>%
  bg(., i = ~`Difference (%)` >= 20 & `Difference (%)` < 30, part = "body", bg = "yellow") %>%
  bg(., i = ~`Difference (%)` >= 30 | NHSE == 0, part = "body", bg = "red")
  
hivtests_tbl
```

# HBV diagnoses

```{r hbv_diags, echo=FALSE}
hbvdiags <- sentinel_nhse_compared %>%
filter(event == "HBV test") %>%
select(-event) %>%
rename(Site = site_code,
NHSE = nhse_total,
ECDS = ecds_total)

hbvdiags$perc_diff <- as.numeric(hbvdiags$perc_diff)
                                    
hbvdiags <- hbvdiags %>%
  mutate(perc_diff = round(perc_diff, 1)) %>%
  rename(`Difference (%)` = perc_diff,
         `Difference (absolute)` = abs_diff)
       
# define values
mean_nhse <- round(mean(hbvdiags$NHSE), 0)
total_nhse <- round(sum(hbvdiags$NHSE), 0)
mean_ecds <- round(mean(hbvdiags$ECDS), 0)
total_ecds <- round(sum(hbvdiags$ECDS), 0)
mean_diff <- round(mean(hbvdiags$perc_diff), 0)
max_diff <- hbvdiags[which.max(hbvdiags$perc_diff),]
min_diff <- hbvdiags[which.min(hbvdiags$perc_diff),]
```

### Fig.7a: Difference (%) in number of ECDS ED HBV diagnoses compared to NHSE data
```{r hbvdiag_graph, echo=FALSE, fig.width=7, fig.height=5}
ggplot(hbvdiags, aes(x = reorder(Site, -`Difference (%)`), y = `Difference (%)`)) +
geom_col(aes(fill = `Difference (%)`), position = "dodge2") +
scale_y_continuous(labels = label_comma(), expand = c(0, 0)) +
scale_fill_continuous(trans = 'reverse') +
labs(x = "Site") +
theme_ukhsa() +
theme(legend.position = "none",
  text = element_text(size=12),
  legend.text = element_text(size=12),
  axis.text.x = element_text(angle = 90),
  plot.subtitle = element_text(vjust = 5))
```

### Fig.7b: Difference (absolute) in number of ECDS ED HBV diagnoses compared to NHSE data
```{r hbvdiag_graph2, echo=FALSE, fig.width=7, fig.height=5}
ggplot(hbvdiags, aes(x = reorder(Site, -`Difference (absolute)`), y = `Difference (absolute)`)) +
geom_col(aes(fill = `Difference (absolute)`), position = "dodge2") +
scale_y_continuous(labels = label_comma(), expand = c(0, 0)) +
scale_fill_continuous(trans = 'reverse') +
labs(x = "Site") +
theme_ukhsa() +
theme(legend.position = "none",
  text = element_text(size=12),
  legend.text = element_text(size=12),
  axis.text.x = element_text(angle = 90),
  plot.subtitle = element_text(vjust = 5))
```

### Table 6: Numbers of ED attendances with HBV diagnoses in NHSE and ECDS data by site with percentage difference (ECDS compared to NHSE)
```{r hbvdiags_table, echo=FALSE}
# green if less than 20%
# yellow if 21-30%
# red if over 30%
hbvdiags_tbl <- hbvdiags %>%
  arrange(`Difference (%)`) %>%
  flextable() %>%
  autofit() %>%
  bg(., i = ~`Difference (%)` < 20, part = "body", bg = "green") %>%
  bg(., i = ~`Difference (%)` >= 20 & `Difference (%)` < 30, part = "body", bg = "yellow") %>%
  bg(., i = ~`Difference (%)` >= 30| NHSE == 0, part = "body", bg = "red")
  
hbvdiags_tbl
```

# HCV diagnoses

```{r hcv_diags, echo=FALSE}
hcvdiags <- sentinel_nhse_compared %>%
filter(event == "HCV diagnosis") %>%
select(-event) %>%
rename(Site = site_code,
NHSE = nhse_total,
ECDS = ecds_total)

hcvdiags$perc_diff <- as.numeric(hcvdiags$perc_diff)
                                    
hcvdiags <- hcvdiags %>%
  mutate(perc_diff = round(perc_diff, 1)) %>%
  rename(`Difference (%)` = perc_diff,
         `Difference (absolute)` = abs_diff)
       
# define values
mean_nhse <- round(mean(hcvdiags$NHSE), 0)
total_nhse <- round(sum(hcvdiags$NHSE), 0)
mean_ecds <- round(mean(hcvdiags$ECDS), 0)
total_ecds <- round(sum(hcvdiags$ECDS), 0)
mean_diff <- round(mean(hcvdiags$perc_diff), 0)
max_diff <- hcvdiags[which.max(hcvdiags$perc_diff),]
min_diff <- hcvdiags[which.min(hcvdiags$perc_diff),]
```

### Fig.8a: Difference (%) in number of ECDS ED HCV diagnoses compared to NHSE data
```{r hcvdiag_graph, echo=FALSE, fig.width=7, fig.height=5}
ggplot(hcvdiags, aes(x = reorder(Site, -`Difference (%)`), y = `Difference (%)`)) +
geom_col(aes(fill = `Difference (%)`), position = "dodge2") +
scale_y_continuous(labels = label_comma(), expand = c(0, 0)) +
scale_fill_continuous(trans = 'reverse') +
labs(x = "Site") +
theme_ukhsa() +
theme(legend.position = "none",
  text = element_text(size=12),
  legend.text = element_text(size=12),
  axis.text.x = element_text(angle = 90),
  plot.subtitle = element_text(vjust = 5))
```

### Fig.8b: Difference (absolute) in number of ECDS ED HCV diagnoses compared to NHSE data
```{r hcvdiag_graph2, echo=FALSE, fig.width=7, fig.height=5}
ggplot(hcvdiags, aes(x = reorder(Site, -`Difference (absolute)`), y = `Difference (absolute)`)) +
geom_col(aes(fill = `Difference (absolute)`), position = "dodge2") +
scale_y_continuous(labels = label_comma(), expand = c(0, 0)) +
scale_fill_continuous(trans = 'reverse') +
labs(x = "Site") +
theme_ukhsa() +
theme(legend.position = "none",
  text = element_text(size=12),
  legend.text = element_text(size=12),
  axis.text.x = element_text(angle = 90),
  plot.subtitle = element_text(vjust = 5))
```

### Table 7: Numbers of ED attendances with HCV diagnoses in NHSE and ECDS data by site with percentage difference (ECDS compared to NHSE)
```{r hcvdiags_table, echo=FALSE}
# green if less than 20%
# yellow if 21-30%
# red if over 30%
hcvdiags_tbl <- hcvdiags %>%
  arrange(`Difference (%)`) %>%
  flextable() %>%
  autofit() %>%
  bg(., i = ~`Difference (%)` < 20, part = "body", bg = "green") %>%
  bg(., i = ~`Difference (%)` >= 20 & `Difference (%)` < 30, part = "body", bg = "yellow") %>%
  bg(., i = ~`Difference (%)` >= 30| NHSE == 0, part = "body", bg = "red")
  
hcvdiags_tbl
```

# All data
```{r all_data_prep, echo=FALSE}
all_data <- sentinel_nhse_compared 

all_data$perc_diff <- as.numeric(all_data$perc_diff)

all_data <- all_data %>%
  mutate(perc_diff = round(perc_diff, 0)) %>%
  mutate(perc_diff = ifelse(is.na(perc_diff), "No data", perc_diff)) %>%
  select(-c(nhse_total, ecds_total, abs_diff)) %>%
  spread(event, perc_diff) %>%
  rename(Site = site_code) %>%
  select(Site, Attendance, `Blood test`, `HBV test`, `HCV test`, `HIV test`, `HBV diagnosis`, `HCV diagnosis`)
```

### Table 8: Numbers of ED attendances in NHSE and ECDS data by event and site with percentage difference (ECDS compared to NHSE)
```{r all_table, echo=FALSE}
# green if less than 20%
# yellow if 21-30%
# red if over 30%

all_tbl <- all_data %>%
  arrange(Site) %>%
  flextable() %>%
  autofit() %>%
  add_header_row(values = c("", "Percentage change (%)"), colwidths = c(1, 7)) %>%
  flextable::align(i = NULL, j = NULL, align = "center", part = "all")

all_tbl
```