---
title: "Record Linkage Analysis"
date: "Feb 2024"
output: 
  html_document:
    number_sections: yes
    toc: yes
    theme: cerulean 
pdf_document:
  toc: yes
---

------------------------------------------------------------------------

# Setup

## Import R libraries

```{r packages, message=FALSE}
# Import libraries
pacman::p_load(
  pacman,
  tidyverse,
  readxl,
  writexl,
  DBI,
  odbc,
  knitr,
  data.table,
  kableExtra,
  janitor,
  report,
  naniar
)
```

## Import and clean data

```{r dataclean}
source("data_preparation.R")
```

## View data completeness and uniqueness
```{r primary_summary_view}
primary_data_summary
```

```{r secondary_summary_view}
secondary_data_summary
```

- Record identifier is complete for `r prop_complete(primary_clean_data$record_id)`% of records in Primary dataset and `r prop_complete(secondary_clean_data$record_id)`% of records in Secondary dataset. This will facilitate 'Deterministic Linkage'.

- Assumption for this exercise is that the record identifier is unique and correct.

- We will link both datasets using available identifiers to create subset that contains 'comparison pairs'.

- This subset can be used as a 'Gold-standard' from which matching probabilities can be calculated.

## Naming conventions

- Note that both datasets contain variables with the same naming convention, which will cause confusion when identifying these variables after linkage.

- Therefore, one approach to overcome this, is to add a prefix to each variable:

    - For variables in Primary file add P_
    - For variables in Linking file add L_
    
- The names have been formatted like this in primary_clean and secondary_clean

# Deterministic linkage with record identifier

- Perform a cross-join on the identifier, where P_id == L_id.
- Reorder columns & sort by ascending ids

```{r}
rm(list=setdiff(ls(), c("primary_clean", "secondary_clean"))) # removes all unused objects to save memory

# Merge primary and linking based on the 'id' variable
merged_df <- merge(secondary_clean, primary_clean, 
                  by.x = "L_record_id", by.y = "P_record_id", 
                  incomparables = NA) %>%
  select("P_id", 
         "L_id", 
         "L_record_id", 
         "P_birth_date", 
         "L_birth_date",
         "P_location",
         "L_location",
         "P_category", 
         "L_category",
         "P_date",
         "L_date")

# Sort
sorted_indices <- order(merged_df$P_id, merged_df$L_id)
merged_df <- merged_df[sorted_indices, ]
rm(sorted_indices)
```

[Continue with detailed data processing steps...]

## Create Agreement patterns

```{r}
# New variables showing agreement patterns 
merged_df$agree_birth <- if_else(merged_df$P_birth_date == merged_df$L_birth_date, 1,0,0)
merged_df$agree_location <- if_else(merged_df$P_location == merged_df$L_location, 1,0,0)
merged_df$agree_category <- if_else(merged_df$P_category == merged_df$L_category, 1,0,0)
merged_df$agree_date <- if_else(merged_df$P_date == merged_df$L_date, 1,0,0)

merged_df$agreement_pattern <- paste(merged_df$agree_birth, 
                                   merged_df$agree_location, 
                                   merged_df$agree_category, 
                                   merged_df$agree_date, sep = "")

# Generate m-probabilities and u-probabilities
[... continue with probability calculations ...]

# Calculate weights and generate match scores
[... continue with weight calculations ...]

# Final match selection
[... continue with match selection logic ...]
```

## Results Summary

```{r}
# Generate final summary statistics
summary_stats <- data.frame(
  Metric = c("Total Records Processed",
             "Deterministic Matches",
             "Probabilistic Matches",
             "Total Matched Pairs"),
  Count = c(nrow(primary_clean),
            sum(deterministic_matches),
            sum(probabilistic_matches),
            total_matches)
)

kable(summary_stats, caption = 'Linkage Summary') %>%
  kable_styling(bootstrap_options = c("striped"), 
                full_width = F, 
                position = "left")
```

## Visualization of Results

```{r}
# Generate visualizations of match quality
ggplot(match_quality, aes(x = weight_score)) +
  geom_histogram(fill = "lightblue", bins = 30) +
  labs(title = "Distribution of Match Weights",
       x = "Weight Score",
       y = "Frequency") +
  theme_minimal()
```

# Complete